---
title: "Using The headhoncho Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using The headhoncho Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
cat("\n", getwd(), "\n")
```


```{r, include = FALSE}
options(gargle_oauth_cache = "../.secrets")
if (dir.exists("../.secrets")) {
  googlesheets4::gs4_auth(cache = "../.secrets", email = "*@smith.edu")
  can_authenticate <- TRUE
} else {
  can_authenticate <- FALSE
}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = can_authenticate
)
```

```{r setup}
library(headhoncho)
```


This article is designed give a brief overview of the functionality of the `xkcd()` package. 

## The `read_google_sheet()` function

This function is the `headhoncho` equivalent to the `googlesheets4::read_sheet()` function. By taking in an identifier of a (private, in your Google Drive account) Google Sheet, this function converts the data into a tibble that can be stroed in your environment and manipulated. This is the reccomended first step, although you are able to skip over it if your data of interest is originally stored in a non-Google-Sheet file. 

```{r}
example_data <- read_google_sheet("1HO3e_VCqdbyRGbKOgzt57aLLlhVCP7YQGmy411tqGNc")
example_data
```


## The `split_sheet()` function

This function is the main functionality of the package. It takes in a dataframe object (typically a tibble that has been `read_google_sheet()`) as well as the columns on which to group and split. It outputs a list of dataframes, where each dataframe contains a single group of the splitting variables. Columns should be specified using the `data_frame$variable` format, and as many columns as desired can be specified as individual arguments (no need to group them manually into any sort of list to pass to the function!). After creating the list of dataframes, this function automatically discards any dataframes for which there are 0 observations, but this behavior can be overridden by specifying `keep = TRUE`. For example, here is a dataframe in which observations will be discarded:

```{r}
df <- data.frame(
  A = c("a", "a", "b", "b"),
  B = c(1, 2, 2, 3),
  C = c(TRUE, TRUE, TRUE, FALSE)
)

## Don't keep data frames with no observations (ex: group formed by df$A == "a" and df$B == 3)
data_list <- split_sheet(df, df$A, df$B)
data_list

## Keep all data frames regardless of if they are empty.
data_list_2 <- split_sheet(df, df$A, df$B, keep = TRUE)
data_list_2
```

Here is the `split_sheet()` function when used on the headhoncho mock dataset:

```{r}
example_data_list <- split_sheet(example_data, example_data$Citizenship_DESC)
example_data_list
```


## The `export_data()` function

The `export_data()` function does the opposite of the `read_sheet()` function: it takes a (list of) dataframe(s) in your R environment and exports them to separate Google Sheets. This function will default to creating new Google Sheets to be written, but can take a vector of sheet identifiers if there are specific sheets that you would like to overwrite. 


```{r, eval = FALSE}
#Automatically will create new sheets and start importing data at the first worksheet within them,
#but that can be overidden if desired Setting the names of these spreadsheets to reflect their
#contents. Since 'Domestic Students' comes first in the data list made earlier, put that one first.
export_data(data_list, names = c("Domestic Students", "International Students"))
```


